\chapter{Distribution of Tasks}
\label{app:distribution}

This appendix describes the distribution of tasks between the developers of this Proof of Concept (PoC) implementation of the PSS Integration Framework.

\section*{Research}

The research necessary for the development of the PSS-IF PoC was done by both, Bernhard Radke and Konstantin Govedarski.

\section*{Conceptual Development}

The conceptual development of the PSS-IF PoC was done by both, Bernhard Radke and Konstantin Govedarski. Both authors are in comparable knowledge of the concepts and technology behind each component of the PoC system.

\section*{Implementation}

The implementation of the PSS-IF PoC was made by both, Bernhard Radke and Konstantin Govedarski. The approximate distribution of specific tasks is provided below.

\begin{itemize}
\item \textbf{Utilities}: \textit{Bernhard Radke} and \textit{Konstantin Govedarski}
\item \textbf{Core}:
	\begin{itemize}
	\item \textbf{Metamodel}:
		\begin{itemize}
		\item \textbf{Node and Edge Types}: \textit{Bernhard Radke} and \textit{Konstantin Govedarski}
		\item \textbf{Connection Mappings}: \textit{Bernhard Radke}
		\item \textbf{Attributes}: \textit{Konstantin Govedarski}
		\item \textbf{Inheritance}: \textit{Bernhard Radke}
		\end{itemize}
	\item \textbf{Model}: \textit{Bernhard Radke} and \textit{Konstantin Govedarski}
	\item \textbf{Operational API}: \textit{Bernhard Radke}
	\item \textbf{Canonic Metamodel}: \textit{Konstantin Govedarski}
	\item \textbf{Canonic Persistence}: \textit{Bernhard Radke}
 	\end{itemize}
\item \textbf{Transformations}:
	\begin{itemize}
	\item \textbf{Views}: \textit{Bernhard Radke}
	\item \textbf{Operators}: \textit{Bernhard Radke}
	\item \textbf{API}: \textit{Konstantin Govedarski}
	\item \textbf{Generic Graph}: \textit{Konstantin Govedarski}
	\end{itemize}
\item \textbf{Languages and Serialization}:
	\begin{itemize}
	\item \textbf{GraphML}: \textit{Bernhard Radke}
	\item \textbf{EPC (Visio)}: \textit{Konstantin Govedarski}
	\item \textbf{BPMN (Visio)}: \textit{Konstantin Govedarski}
	\item \textbf{SysML4Mechatronics}: \textit{Konstantin Govedarski}
	\end{itemize}
\end{itemize}

\section*{Documentation}

The documentation of the PSS-IF PoC was created by both, Bernhard Radke and Konstantin Govedarski. The documentation includes:

\begin{itemize}
\item This document.
\item The documentation of the source code (JavaDocs).
\end{itemize}